The resurgence of the linear optics interferometer --- recent advances \& applications

### 2017-05-13 paper.tex
Revised intro

Linear optics play a prominent role in quantum information processing. Photons make fantastic ``flying" qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting \cite{bib:KLM01}. Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

@@

In the last decade, single photons have been produced mainly through a nonlinear optical process know as spontaneous parametric down-conversion (SPDC), a nonlinear optical process. SPDC produces single photons in correlated pairs that has been a staple source of entanglement. Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical labs, and three photons have become the standard benchmark. Experiments using up to ten photons have been demonstrated. We have also seen the meteoric rise of quantum dots in the role for producing single photons. It is now possible to use a quantum dot to produce multiple single photons which are highly indistinguishable. Quantum dots produce single photons on demand, a significant advantage over SPDC which is an inherently probabilistic process.

Another approach to linear optics is to pack more features known as degrees of freedom onto a single photon, and to increase the degree of control over them. It is possible to manipulate degrees of freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Such methods open up new ways to engineer quantum states and quantum operations, and expand the toolbox of baseline resources we have for implementing quantum computations, which could prove crucial in the race to achieve quantum supremacy.

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction}.

@@

\section{State preparation}

@@

\section{Photodetection}

>>>>

Linear optics play a prominent role in quantum information processing. Photons make fantastic ``flying" qubits, and are readily used for quantum communication \cite{bib:Northup14} and quantum key distribution \cite{bib:Lo14}. In 2001, Knill, Laflamme and Milburn (KLM) showed that efficient quantum computing is possible using only linear optical components, that is single photons, beamsplitters, phase shifters and photon counting \cite{bib:KLM01}. Furthermore, an optical implementation of quantum protocols can potentially piggyback on existing communication infrastructure. Combined with the relative ease of transporting quantum states of light, as compared to say trapped ions or superconducting qubits, linear optics forms a powerful platform for quantum information processing.

@@

In the last decade, single photons have been produced mainly through a nonlinear optical process know as spontaneous parametric down-conversion (SPDC), a nonlinear optical process. SPDC produces single photons in correlated pairs that has been a staple source of entanglement. Through developments in production, the number of high quality indistinguishable photons that can be simultaneously produced have increased steadily over the years. Single-photon pairs, once a novelty, are now run-of-the-mill business for quantum optical labs, and three or four photons have become commonplace \cite{bib:Spagnolo12, bib:Tillmann13, bib:Spring13,bib:Broome13}. Experiments using up to ten photons have been demonstrated \cite{bib:WangChen16, bib:Chen17}. We have also seen the meteoric rise of quantum dots in the role for producing single photons \cite{bib:Ding16,bib:WangHe16, bib:213601}. They are a scalable source of single photons which are highly indistinguishable, and are produced on demand--a significant advantage over SPDC which is an inherently probabilistic process.

Another approach to linear optics is to pack more features known as degrees of freedom onto a single photon, and to increase the degree of control over them. It is possible to manipulate degrees of freedom like polarization, time-of-arrival, and orbital angular momentum \cite{bib:Tillmann2015,bib:Bozinovic2013, bib:Nicolas2014, bib:Humphreys2013, bib:Donohue2013}. Since KLM, new proposals for linear optical quantum computations (LOQC) have surfaced \cite{bib:Nielsen04,bib:BrowneRudolph05,bib:Gimeno-Segovia15,bib:Pant17}. Others do not promise universality in their processing, but still performs tasks such as sampling \cite{bib:Aaronson11}, and secure quantum computations \cite{bib:Tan16} that might be hard or impossible with a classical system.  Such methods open up new ways to engineer quantum states and quantum operations, and expand the toolbox of baseline resources we have for implementing quantum computations, which could prove crucial in the race to achieve quantum supremacy.

We begin in Section \ref{sec:math} with an overview of the mathematical treatment of linear optical transformation on single photons, and in Section \ref{sec:encoding} with a description of some commonly used qubit encodings. Any unitary transformation acting on the spatial label of single photons can be efficiently decomposed into a network of beamsplitters and phase-shifters. We discuss this, and an extension to handle additional internal states, in Section \ref{sec:decomp}. For applying any linear optical circuit, one needs an accurate description of its unitary representation. A full quantum tomography is costly and time-consuming. Some practical methods have been devised to do this reconstruction without resorting to tomography, and these are described in Section \ref{sec:reconstruction}. In Section \ref{sec:state}, we recap the experimental breakthroughs in state preparation of single photons, and some entangled quantum states. This is followed by advances in linear optical networks, and photodectors in Sections \ref{sec:network} and \ref{sec:detector} respectively. Last, but not least, we review applications for linear optical interferometry in LOQC, {\sc bosonsampling}, and quantum metrology.

@@

\section{State preparation}\label{sec:state}

@@

\section{Photodetection}\label{sec:detector}
